---
title: "How to connect to motherduck"
format: html
eval: false
vignette: >
  %\VignetteIndexEntry{How to Connect to MotherDuck}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: load
#| echo: false
#| message: false
#| warning: false
#| error: false
#| include: false


hook <- fansi::set_knit_hooks(knitr::knit_hooks, which = c("output", "warning", "error", "message"))

options(crayon.enabled = TRUE)


```



### Create a duckdb instance and connect to your motherduck account

When creating a duckdb database, you have three options

1. A in-memory based instance that exists in your local computer
2. A file based instance that exists in your local computer
3. A cloud-based instance through [motherduck](https://motherduck.com/)

To create options 1 or 2 you can simply use either the [duckdb](https://duckdb.org/) or the [duckplyr](https://duckplyr.tidyverse.org/index.html) packages.

To use option 3, you will need to create a motherudck account and generate an access token. Once created, save your access token to an your R enviorment with `usethis::edit_r_environ()`. I recommend using  `MOTHERDUCK_TOKEN` as your variable name.

Once completed, you can simply use the `connect_to_motherduck()` function and pass through your token variable name and optional configuration options.



>### Whats the difference between Motherduck and Duckdb?
>
> - Duckdb is a database that you can deploy and run either temporary or permanently in your computer. If you run it via your local computer, it is only available on your computer
>
- Motherduck is a cloud based deployment of duckdb which means you can save your data in the cloud or access it locally
>
- Most core functions in this package work for both motherduck or duckdb database
>
- It is more of a question if you want you data to be access only locally on your computer or if you want to be able to access it remotely via the cloud

```{r}
#| label: connection-example
#| echo: true
#| warning: true
#| message: true
#| include: true



con_md <- connect_to_motherduck("MOTHERDUCK_TOKEN")

```

This will return a connection and print statement indicating if connection status.

At any time you can validate your connection status with `validate_md_connection_status()`


```{r}
#| label: validate-examples
#| echo: true
#| warning: true
#| message: true
#| include: true

validate_md_connection_status(con_md)

```



### How to create a motherduck account and access token?

1. Go to [motherduck](https://motherduck.com/) and create an account, free options are available
2. Go to your user name in the top right, click settings then click access tokens
3. Click create token and then name your token and copy the token code
4. You will need this token to access your account 
5. If you want to access it via R then simplest way is to save your access code as a variable in your r environment 
6. Simply leverage the [usethis](https://usethis.r-lib.org/) function `usethis::edit_r_environ()` to set your access code to a variable and save it -- this is one time activity
7. To check if you have correctly saved your variable then you can use the Sys.getenv("var_name") with substituting "var_name" with your access token name
8. Going forward, if you want to access your token you don't need to re-type the access token, simply remember your variable name

### How to configure our connection?

When connecting to motherduck there are a number of configuration options available, you can reference them via the `motherduck::config_db` which will pull a list of options and their default values

To change these, simply edit the configuration options you want and then pass the list as an argument `connect_to_motherduck()` or `duckdb::duckdb()` if connecting locally

You can see the full list of duckdb configuration options [here](https://duckdb.org/docs/stable/configuration/overview.html) or alternatively you can use `list_setting()` to see your current configuration options.

```{r}
#| label: connection-config
#| echo: true
#| warning: true
#| message: true
#| include: true


config <- motherduck::config_db 


config$allow_community_extensions <- "true" 

con_md <- connect_to_motherduck("MOTHERDUCK_TOKEN",config = config) 

```

- Access the default configuration options via the built in list
- Change the options you want, ensure you leverage duckdb syntax
- Pass the modtified list to config option in the function

At any time you can see what configuration arguments are for your connection with `list_setting()`.

```{r}
#| label: list-settings
#| echo: true
#| warning: false
#| message: true
#| include: true


list_setting(con_md)

```



```{r}
#| label: list-settings-hidden
#| echo: false
#| warning: false
#| message: true
#| include: true


list_setting(con_md) |> 
    head(10) 

```
Congratulations, you've set connected to your motherduck database from R!

Please see the {motherduck} package down website to see additional documentation on how to use the functions and motherduck
